A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-13 main.tex
Add comments to the section on staged semantics.

The reduction given above is a full reduction and any redexes can be reduced in arbitrary order.
But we need to fix order of reduction when we implement an interpreter of \LMD and
the order must be appropriate in terms of stages.

We will define deterministic call-by-value staged semantics which can be used as a basis of a \LMD\ interpreter.
In this reduction, $\longrightarrow_\beta$ and $\longrightarrow_\Lambda$ are allowed only at stage $\epsilon$ and 
$\longrightarrow_\blacklozenge$ is allowed only at stage $\alpha$.

\begin{definition}[Values]
	The family $V^A$ of sets of values, ranged over by $v^A$ and 
	the sets of $\epsilon$-redexes (ranged over by $R^\epsilon$) and $\alpha$-redexes (ranged over by $R^\alpha$)
	are defined by following grammar. In the grammar $A \neq \epsilon$.
	\begin{align*}
		\textrm{Values}  &  & v^\epsilon \in V^\epsilon & ::= \lambda x:\tau.M \mid\ \TB_\alpha v^\alpha \mid \Lambda\alpha.v^\epsilon                                       & \\
		                 &  & v^A \in V^A               & ::= x \mid \lambda x:\tau.v^A \mid v^A\ v^A \mid\ \TB_\alpha v^{A\alpha} \mid \Lambda\alpha.v^A \mid v^A\ \epsilon & \\
		                 &  &                           & \quad\   \mid\ \TBL_\alpha v^{A'} (\text{if } A'\alpha = A \text{ and } A' \neq \epsilon)                          & \\
		                 &  &                           & \quad\   \mid \%_\alpha v^{A'} (\text{if } A'\alpha = A)                                                           & \\
		\textrm{Redexes} &  & R^\epsilon                & ::= (\lambda x:\tau.M)\ v^\epsilon \mid (\Lambda\alpha.v^\epsilon)\ \epsilon                                       & \\
		                 &  & R^\alpha                  & ::=\ \TBL_\alpha \TB_\alpha M                                                                                      & \\
	\end{align*}
\end{definition}

Values at $\epsilon$ stage are a $\lambda$ abstraction, a quoted code, or a $\Lambda$ abstraction.
The body of a $\lambda$ abstraction can be any term but the body of $\Lambda$ abstraction only can be a value.
This restriction means that the body of $Lambda$ abstraction must be evaluated before application of an $\epsilon$ stage.
Values at non-$\epsilon$ stage includes all term except $\TBL_\alpha v^\epsilon$.
This is because $\TBL_\alpha v^\epsilon$ is a redex.

\begin{definition}[Evaluation Context]
	The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$, is defined by the following grammar,
	$A$ is nonempty. (Although $B,A'$ can be empty.)
	$A \neq \epsilon$\\
	\begin{align*}
		E^\epsilon_B \in ECtx^\epsilon_B & ::= \square\ (\text{if\ } B = \epsilon) \mid E^\epsilon_B\ M \mid v^e\ E^\epsilon_B     
		\mid \TB_\alpha E^\alpha_B \mid \Lambda\alpha.E^\epsilon_B
		\mid E^\epsilon_B\ \epsilon                                                                                                 \\
		E^A_B \in ECtx^A_B               & ::= \square\ (\text{if } A = B) \mid \lambda x:\tau.E^A_B \mid E^A_B\ M \mid v^A\ E^A_B 
		\mid E^\epsilon_B \mid \TB_\alpha E^{A\alpha}_B
		\mid \TBL_\alpha E^{A'}_B \ (\text{where } A'\alpha = A)                                                                    \\
		                                 & \quad \mid \Lambda\alpha.E^\epsilon_B                                                   
		\mid E^A_B\ \epsilon \mid \%_\alpha\ E^{A'}_B \ (\text{where } A'\alpha = A)                                                \\
	\end{align*}
\end{definition}

We write $E^A_B[M]$ for a term obtained by filling the hole in $E^A_B$ with $M$.

\begin{definition}[Staged Reduction]
	The staged reduction relation, written $M \longrightarrow_s M'$, is defined by 
	the least relation closed under the rules in Figure \ref{fig:staged-reduction}.
\end{definition}

\begin{figure}
	\begin{center}
		\begin{align*}
			E^A_\epsilon [(\lambda x:\tau.M)\ v^\epsilon]       & \longrightarrow_s E^A_\epsilon[M[x\mapsto v^\epsilon]]             \\
			E^A_\epsilon [(\Lambda\alpha.v^\epsilon)\ \epsilon] & \longrightarrow_s E^A_\epsilon[v^\epsilon[\alpha\mapsto \epsilon]] \\
			E^A_\alpha [\TBL_\alpha \TB_\alpha v^\alpha]        & \longrightarrow_s E^A_\alpha[v^\alpha]                             \\
		\end{align*}
		\caption{Staged Reduction}
		\label{fig:staged-reduction}
	\end{center}
\end{figure}

>>>>

The reduction given above is full reduction and any redexes---even
under quotation---can be reduced in an arbitrary order.  Following
previous work~\cite{}, we introduce (small-step, call-by-value) staged
semantics, where only redexes at stage $\varepsilon$\AI{not quite true
  because of unquote--quote reduction} are reduced, modeling an
implementation.

We start with the definition of values.  Since terms under quotations are
not executed, the grammar is indexed by stages.

\begin{definition}[Values]
	The family $V^A$ of sets of values, ranged over by $v^A$,
	is defined by following grammar.  In the grammar, $A \neq \epsilon$ is assumed.
	\begin{align*}
		v^\epsilon \in V^\epsilon & ::= \lambda x:\tau.M \mid\ \TB_\alpha v^\alpha \mid \Lambda\alpha.v^\epsilon                                       & \\
		v^A \in V^A               & ::= x \mid \lambda x:\tau.v^A \mid v^A\ v^A \mid\ \TB_\alpha v^{A\alpha} \mid \Lambda\alpha.v^A \mid v^A\ \epsilon & \\
		                         & \quad\   \mid\ \TBL_\alpha v^{A'} (\text{if } A = A'\alpha \text{ for some } \alpha, A' \neq \epsilon)                          & \\
                                          & \quad\   \mid\ \%_\alpha v^{A'} (\text{if } A = A'\alpha  \text{ for some } \alpha, A')
	\end{align*}
\end{definition}

Values at $\epsilon$ stage are a $\lambda$-abstraction, a quoted code,
or a $\Lambda$ abstraction.  The body of a $\lambda$-abstraction can
be any term but the body of $\Lambda$-abstraction has to a value.  It
means that the body of $\Lambda$-abstraction must be evaluated.  The
side condition for $\TBL_\alpha v^{A'}$ means that escapes in a value
can appear only under nested quotations (because an escape under a
single quotation will splice the code value into the surrounding
code).  See Hanada and Igarashi~\cite{} for details.

\begin{definition}[Evaluation Context and Redex]
  The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$,
the sets of $\epsilon$-redexes (ranged over by $R^\epsilon$) and $\alpha$-redexes (ranged over by $R^\alpha$) are defined by the following grammar.
	In the grammar, $A$ is assumed to be nonempty (but $B,A'$ can be empty).
	\begin{align*}
		E^\epsilon_B \in ECtx^\epsilon_B & ::= \square\ (\text{if\ } B = \epsilon) \mid E^\epsilon_B\ M \mid v^e\ E^\epsilon_B     
		\mid \TB_\alpha E^\alpha_B \mid \Lambda\alpha.E^\epsilon_B
		\mid E^\epsilon_B\ \epsilon                                                                                                 \\
		E^A_B \in ECtx^A_B               & ::= \square\ (\text{if } A = B) \mid \lambda x:\tau.E^A_B \mid E^A_B\ M \mid v^A\ E^A_B 
		\mid E^\epsilon_B \mid \TB_\alpha E^{A\alpha}_B
		\mid \TBL_\alpha E^{A'}_B \ (\text{where } A'\alpha = A)                                                                    \\
		                                 & \quad \mid \Lambda\alpha.E^\epsilon_B                                                   
		\mid E^A_B\ \epsilon \mid \%_\alpha\ E^{A'}_B \ (\text{where } A'\alpha = A)                                                \\
\\
		 R^\epsilon  & ::= (\lambda x:\tau.M)\ v^\epsilon \mid (\Lambda\alpha.v^\epsilon)\ \epsilon                                       & \\
		 R^\alpha    & ::=\ \TBL_\alpha \TB_\alpha M                                                                                      & \\	\end{align*}
\end{definition}

\AI{some more explanation of the syntax of evaluation contexts.}

We write $E^A_B[M]$ for the term obtained by filling the hole in $E^A_B$ by $M$
and define the staged reduction relation.

\begin{definition}[Staged Reduction]
	The staged reduction relation, written $M \longrightarrow_s M'$, is defined by 
	the least relation closed under the rules below.
% \begin{figure}
% 	\begin{center}
		\begin{align*}
			E^A_\epsilon [(\lambda x:\tau.M)\ v^\epsilon]       & \longrightarrow_s E^A_\epsilon[M[x\mapsto v^\epsilon]]             \\
			E^A_\epsilon [(\Lambda\alpha.v^\epsilon)\ \epsilon] & \longrightarrow_s E^A_\epsilon[v^\epsilon[\alpha\mapsto \epsilon]] \\
			E^A_\alpha [\TBL_\alpha \TB_\alpha v^\alpha]        & \longrightarrow_s E^A_\alpha[v^\alpha]                             \\
		\end{align*}
     %    	\caption{Staged Reduction}
     %    	\label{fig:staged-reduction}
     %    \end{center}
     % \end{figure}
\end{definition}
